supplementary material for ``manifold fitting by ridge estimation: a subspace-constrained approach''

### 2021-02-25 rebuttal/response-referee1.tex

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{color}
\usepackage{url}
\usepackage{amsmath,graphicx,amssymb}
\bibpunct{(}{)}{;}{a}{}{,}
%\newcommand{\ignore}[1]{}

%\renewcommand{\baselinestretch}{1.7}
\setlength{\textheight}{9in}
\setlength{\textwidth}{6in}
\addtolength{\hoffset}{-0.45in}
\addtolength{\voffset}{-0.425in}

\newtheorem{thm}{Theorem}[section] %(If you want theorem numbered
\newtheorem{lemma}{Lemma}[section] %%    with section number.
\newtheorem{cor}{Corollary}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{theorem}{Theorem}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\eps}{\epsilon}
\newcommand{\rarr}{\rightarrow}
\renewcommand{\b}{\mathbf b}
\newcommand{\qed}{{\unskip\nobreak\hfil\penalty50\hskip2em\vadjust{}
            \nobreak\hfil$\Box$\parfillskip=0pt\finalhyphendemerits=0\par}}
\newcommand{\ignore}[1]{}
\renewcommand{\thefootnote}{}

\newcommand{\yf}[1]{{\textcolor{red}{[Fan: #1]}}}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%
%%%Last change on May 22 by Jin%%%%
%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\bf \large Point-by-Point Response to Referee One}
\end{center}
Thank you very much for your careful reading our paper as well as your invaluable comments. We list our response to your comments as follows:\\ 
major comments\\

\noindent ``\textit {This paper is on manifold learning.  It is very confusing to read.''}\\
Answer:\\
To the best of our knowledge, manifold learning refers to an approach to non-linear dimensionality reduction, which is a method trying to find the new representation in a lower-dimensional space by approximately keeping the geometry properties of the input which resides in a high-dimensional ambient space. In this paper, we learn the original manifold in the ambient space. Therefore, our work is not on manifold learning.

\noindent ``\textit{For example, realistically the data generated by the manifold model are generated with noise, and the distribution and properties of the noise in conjunction with the reach of the manifold are critical for the recovery. However, these critical issues are bypassed by making strong assumptions whereby only the local behavior of ridges matters.''}\\
Answer:\\
We agree with the referee that the recovery is highly related with the noise and reach of the manifold. In our work, our model in Equation (1) on page assumes the data are generated as 
\begin{equation}\label{model}
x_i=\tilde{x}_i+\epsilon_i,
\end{equation}
where $\tilde{x}_i$ is the noiseless signal which belong to some manifold, and $\epsilon_i$ is the noise. We use the observations $x_i$ to recover the manifold. Therefore, we did not bypass consideration of the noise in our model.
\\
The reach is a global concept which describes the degree of curvature for the manifold. A larger {``\it reach''} indicates the samples resides further from the manifold can also has a unique projection onto $\cal M$. Therefore, the manifold is more flat globally.

\noindent ``\textit{ This approach may be interesting in its own rights, but the impact in the bigger picture is rather limited due to the limited perspective and the strong assumptions that are needed. The best part are the computational perspectives provided in the paper, while the theoretical developments are also not very interesting.'' }\\
Answer:\\
In fact, Assumption 2.2 is more likely to be thought as a ``Fact'' rather than an Assumption. The main intention of Assumption 2.2 is to say that any low-dimensional manifold can be thought as the ridge of some unknown density function. In this new version of the manuscript, we modify the presentation of Assumption 2.2 to avoid the confusion.
\\
\noindent ``\textit {A key problem is that the errors in the sampling of the data generated by the manifold are not modeled and the noise-contaminated data are not studied. The paper is not well written, the presentation is substandard  and is unfocused.'' }\\
Answer:\\
As in Equation \eqref{model}, we model the data as the noiseless $\tilde{x}_i\in \cal M$ plus the noise $\epsilon_i$. We consider the noise in the theories and all the cases of the numerical experiments.\\

\noindent ``\textit{Problematic assumptions include Assumption 2.2, where parts 1 and 3 are unclear and ambiguous and Assumption 3.4 which is equally unmotivated.''}\\
Answer:\\
1. Assumption 2.2 only wants to show that any noiseless manifold can be thought as the ridge of some background density function $p(x)$. And the method to construct $p(x)$ is not complicated.\\
2. Because of $\mu_s(x,r,h)\in(0,1),\mu_{s,t}(x,r,h)\in (0,1)$, we know 
\[
\mu(r,h) = \sup_x \max\{\max_s \mu_s(x,r,h), \max_{s,t}\mu_{s,t}(x,r,h)\}\in [0,1].
\]
Assumption 3.4 is a bit stronger by requiring $\mu(r,h)$ to be upper-bounded by a number $\ell$ which is less than 1 for all $h$.\\
 \noindent ``\textit{The simulations are too limited as there are many other competing methods to recover the manifolds that are considered, and these manifolds are too simple to be of much interest.'' } \\
 Answer:\\
 We compare four manifold fitting algorithms on the one-dimensional ring manifold to show the good recovery property of {\it l}-SCRE under the criteria of average margin and Hausdorff distance. The manifold fitting result does not rely on the dataset too much. The comparison results on the other dataset are similar. To justify, we add another numerical result corresponding to a 2-dimensional sphere in the 3D ambient space.

%\noindent ``\textit {1. My major concern is about the condition needed for the establishment of results in this paper.. The authors need $P^*$ to have constant eigengap, which is far too strong and not realistic (it implies $p_{ij}$ are in the constant order). Under this regime, the results the authors present are expected, not surprising, and more or less have been long known to theoretical researchers working on community detection. This limits the depth and scope of this paper. In community detection, a less trivial regime is when $p_{i,j}=o(1)$. A more interesting case is when $p_{ij}$ is around $\log n/n$. An exciting regime is when $p_{ij}=O(\log n/n)$, where the exact recovery is impossible. I think this paper is a good start and the authors are encouraged to continue along this path.}''\\
%Answer: \\
%Thanks for the important suggestion. We have researched a lot in this boundary of interest. In our new revision, we no longer require $p_{ij}$ are in the constant order, and have relaxed the condition to  $\max_{k_1,k_2}p_{i,j}=\varphi_n n^{-1}$ (see Assumption 4 in the revised paper), where $\varphi_n\to\infty$ and $\varphi_n\le n$ . It covers the regime $p_{ij}=O(\log n/n)$ as a special case. Our models and main theorems in the paper have been modified largely as condition of $p_{ij}$ changed.

%We also highlight that, in the revised paper, we have proved the strongly consistency of the proposed method with a better $l_\infty$-norm bound relative to SCORE and its variants, where the almost surely consistency is not provided there (or in other existing literature sources) and the bound there has been only in $L_F$. Please see the revised paper for details.
%\\
%
%\noindent ``\textit{2. There is much room for this paper to be better presented. There are also some obvious typos in the statements of theorems. The Equation 3.3 should be about $\gamma$ instead of $\lambda$. In Theorem 4.3, $Q^K$ does not need to appear.''}\\
%Answer:\\
%We have corrected them. Meanwhile, we have largely rewritten the paper and the paper went through a proofread process.\\
%
%\noindent ``\textit{3. In the paragraph after the Theorem 4.3, the authors seem to conclude that spectral clustering gives the exact recovery. If so, then maybe it can be stated as a theorem/corollary.''}\\
%Answer:\\
%Yes, you are right. Thank you very much for your suggestions. The consistency of the proposed method is discussed in subsection 4.2.2 where  Theorem 4.4 is added to explain the strong consistency of SCDRE under the challenging region. Moreover, Theorem 4.4 does imply exact recovery. Also, we would like to mention that Theorem 4.4 holds when the community number greater than or equal to 2, without any additional conditions on the edge probability.\\
%
%
%\noindent ``\textit{4. In Section 4.2.2, I am not sure if the truncation step is necessary. The supporting argument the authors give in Section 4.2.1 is not convincing.}''\\
%Answer:\\
%The truncation step on $\hat R_\zeta(i)$ in Section 4.2.1 is indeed no longer needed now and it has been removed in the revised version. In fact, we have realized that this is also one of the advantages of our method comparing to the recent SCORE and its variants. The details can be found in the subsection 4.2.3 which.\\
%
%\noindent ``\textit{5. In the statement of Theorem 4.1, referring $C_n$ a constant associated with $n'$ may cause confuse. There must be better ways to call it.}''\\
%Answer:\\
%For the statement of Theorem 4.1, the estimator of $K$, we have changed the notation $C_n$ in the estimator of $K$, which is a better expression in the regime $p_{ij}=O(\varphi_n n^{-1})$.

%%%%%%%%%%%%
%%%%%%%%%%%%
%%%%%%%%%%%%
%%%%%%%%%%%%
\end{document}

### 2021-02-25 Ridge_Estimation_Supplementary/Ridge-supplementary-final.tex
zzh

\begin{frontmatter}
\title{supplementary material for "manifold fitting by ridge estimation: a subspace-constrained approach"}
%\title{A sample article title with some additional note\thanksref{t1}}
\runtitle{Manifold fitting by ridge estimation}
%\thankstext{T1}{A sample additional note to the title.}

@@

\begin{abstract}
In this supplement we present the technical proofs for the main work. Equation and theorem references made to the main document do not contain letters.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please use \tableofcontents for articles %%
%% with 50 pages and more                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tableofcontents

\begin{appendix}
\section{PROOFS FOR MAIN THEOREMS AND LEMMAS}
In this section, we prove all key theorems and lemmas in the order they appear.
\subsection{Ridge Derivative Lemma}
\begin{lemma}{ 
For any $R_1$, $R_2$, and any point $x_1\in R_1$, the pairwise distance from $x_1$ to $R_2$ yields the order of: %$\min_{x_2\in R_2}\|x_1-x_2\|_2$ %between any two ridges $R_1$ and $R_2$ 
%is of order $O(\|H_1(x_1)-H_2(x_1)\|_F+\|g_1(x_1)-g_2(x_1)\|_2)$, 
\[
\min_{x_2\in R_2}\|x_1-x_2\|_2 = O(\|H_1(x_1)-H_2(x_1)\|_F+\|g_1(x_1)-g_2(x_1)\|_2)
\]
where $H_1(x_1), g_1 (x_1)$ are the Hessian and gradient of some estimated density function $p_1(x_1)$ evaluated at $x_1$; $H_2(x_1)$ and $g_2(x_1)$ are the Hessian and gradient of the density function of $p_2(x)$ evaluated at $x_1$, respectively.
%For any two ridges $R, \hat{R}$ defined by $R = \{x| \Pi_{H}(x)g(x) = 0\}$,$\hat{R} = \{x| \Pi_{\hat{H}}(x))\hat{g}(x) = 0\}$, 
 %$d(x, \hat{R}) = O(\|\hat{H}(x)-H(x)\|_F+\|\hat{g}(x)-g(x)\|_2)$
 }\label{margin}
\end{lemma}

The following proof is a revised simple version of a similar proof in \cite{genovese2014nonparametric}. For completeness, we also include it in our paper.

\begin{proof}\label{Ridge Derivative Lemma}
 For two ridges $R_1, R_2$, we have two density functions $p_1(x)$ and $p_2(x)$ such that the points on each ridge satisfy the solution of
 $\Pi_{H_1}(x) g_1(x) = 0$ 
 and $\Pi_{H_2}(x) g_2(x) = 0$ respectively. For any starting point $x_a \in R_1$, we can build a unit speed curve $\gamma(s)$ derived from the gradient and Hessian of $p_2(x)$ as
 \begin{gather*}
 \gamma_2(0) = x_a \in R_1,\\
 \gamma_2(t_0) = x_b \in R_2,\\
 \gamma'_2(s) = \frac{\Pi_{H_2}(\gamma(s))g_2(\gamma(s))}{\|\Pi_{H_2}(\gamma(s))g_2(\gamma(s))\|_2}.
 \end{gather*}
Note that the curve $\gamma(t)$ connect $x_a$ with $R_2$ by $x_b$. Define the univariate function $\xi(s)$ as
  \[
  \xi_2(s) = p_2(\gamma_2(t_0))-p_2(\gamma_2(s)), \quad 0<s<t_0.
  \]
Through a simple computation, we know 
 \[
 \xi'_2(s)=-\langle  g_2(\gamma_2(s)), \gamma'_2(s) \rangle =- \|\Pi_{H_2}(\gamma_2(s))g_2(\gamma_2(s))\|_2, \quad \xi'_2(t_0) = 0.
 \]
 %\[
 %\xi''(s) = \frac{\langle L_{\gamma'(s)}(\gamma(s)) g(\gamma(s)), L(\gamma(s)g(\gamma(s)))\rangle}{\|L(\gamma(s))g(\gamma(s))\|_2}
 %\]
 %with $\gamma(s)=\frac{L(\gamma(s)g(\gamma(s)))}{\|L(\gamma(s)g(\gamma(s)))\|}$
 The distance from $x_a$ to $R_2$ can be bounded by the curve length of $\gamma_2(t)$ which is $t_0$
 \[
 \begin{aligned}
 &d(x_a, R_2) \\
 =& \|x_a-P_{R_2}(x_a)\|_2\\
 \leq& \|x_a-x_b\|_2 \\
 = &\|\gamma_2(t_0)-\gamma_2(0)\|_2\leq t_0.
 \end{aligned}
 \]
Finally, the problem becomes to bound $t_0$. Suppose $\sup_u \xi''_2(u)>\frac{1}{c}$, by the mean-value theorem, we have
 \[
 \begin{aligned}
 t_0 = &\frac{\xi'_2(t_0)-\xi'_2(0)}{\xi''_2(u)} \\
 = &\frac{\|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2}{\xi''_2(u)} \\
 \leq & c \|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2.
 \end{aligned}
 \]
 Next, we show that $\|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2$ is of the same order with an approximation error of $H_2(x)$ and $g_2(x)$:
 \begin{equation}
 \begin{aligned}\label{bound_temp1}
 &\|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2\\
 =&  \|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0)) - \Pi_{H_1}(\gamma_2(0))g_1(\gamma_2(0))\|_2
 \end{aligned}
 \end{equation}
 Using the triangle-inequality with respect to the $\|\|_2$ norm, we have \eqref{bound_temp1} is dominated by
 \begin{equation}\label{bound_temp2}
 \begin{aligned}
& \|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0)) -\Pi_{H_1}(\gamma_2(0)){g_2}(\gamma_2(0)) \|_2 +...\\
 & \quad \quad  \quad \quad  \quad \quad + \|\Pi_{H_1}(\gamma_2(0)){g_2}(\gamma_2(0)) - \Pi_{H_1}(\gamma_2(0))g_1(\gamma_2(0))\|_2\\
 \end{aligned}
 \end{equation}
 For any matrix $A$, using the Cauchy-Schwartz inequality on each row of $A$, we will get the result $\|Ax\|_2\leq \|A\|_F\|x\|_2$. Thus, similarly, we have \eqref{bound_temp2} is dominated by
 \[
 \begin{aligned}
 & \|\Pi_{H_2}(\gamma_2(0)) - \Pi_{H_1}(\gamma_2(0))\|_F\|g_2(\gamma_2(0))\|_2 +...\\
 &\quad \quad  \quad \quad  \quad \quad  +\|\Pi_{H_1}(\gamma_2(0))\|_F\|g_2(\gamma_2(0))-g_1(\gamma_2(0))\|_2, \\
 \end{aligned}.
 \]
According to the Davis-Kahan theorem, $\|\Pi_{H_2}(\gamma_2(t))-\Pi_{H_1}(\gamma_2(t))\|_F \leq \beta\|H_2(\gamma_2(t))-H_1(\gamma_2(t))\|_F$. The conclusion is proved!
\end{proof}

\subsection{Derivatives' Bias Bound}\label{bias_proof}
\begin{theorem}
The bias of the first order and second order of the $\hat{p}_h(x)$ is 
\begin{equation*}\label{BiasResult}
|{\rm E}(\partial_{x_s}  \hat{p}_h(x)) - \partial_{x_s}p(x)|  = \frac{h^2|\Delta (\partial_{x_s}p(x))|}{2D}  \int \|u\|_2^2 K(u) du+o(h^2),
\end{equation*}
\begin{equation*}\label{order2}
|{\rm E}( \partial_{x_s} \partial_{x_t}  \hat{p}_h(x)) -  \partial_{x_s} \partial_{x_t} p(x)| = \frac{h^2 |\Delta (\partial_{x_s}\partial_{x_t} p(x))|}{2D} \int \|u\|_2^2 K(u) du+o(h^2).
\end{equation*}
where $\Delta$ is the Laplace-Beltrami operator.
\end{theorem}

\begin{proof}
Suppose the kernel function vanishes at infinity for each dimension, i.e., it satisfies $\lim_{u_s\rightarrow \infty}K(u)=0$ for each dimension. Then, using the integration-by-parts formula, we obtain the expectation of first-order derivatives:
\begin{equation}\label{Expectation}
\begin{aligned}
&{\rm E} (\partial_{x_s} \hat{p}_h(x)) \\
\overset{s.1}{=}& \frac{1}{h^D} \int_{y\in \mathbb{R}^D}{\partial_{x_s}}  K(\frac{x-y}{h}) p(y) dy  \\
\overset{s.2}{=}&\frac{1}{h^{D+1}}\int \partial_{z_s} K(z)|_{z=\frac{x-y}{h}} p(y) dy\\
\overset{s.3}{=}&h^{-1} \int_ {u\in \mathbb{R}^D} \partial_{u_s} K(u) p(x-hu)  d u\\
\overset{s.4}{=}&h^{-1}\int_ {u\in \mathbb{R}^D} K(u) \partial_{u_s} p(x-hu) d u \\
\overset{s.5}{=}& \int_ {u\in \mathbb{R}^D} K(u) \partial_{z_s}p(z)|_{z=x-hu} d u.
\end{aligned}
\end{equation}
%We explain the reason for each steps as: 
The equation $s.1$ is the definition of expectation. The equation $s.2$ is obtained from the derivative of the function composition. The equation $s.3$ is obtained from the formula of integration by changing variables. The equation $s.4$ is obtained from the formula of the integration by partition. The equation $s.5$ is similar with that of equation $s.2$.

@@

For the term $u^T H(\partial_{x_s}p(x))u$, we know it is related with the Laplace Beltrami operator of $\Delta(\partial_{x_s}p(x))$ by
\[
\begin{aligned}
\int u^T H(\partial_{x_s}p(x))u K(u) du =& \langle \int uu^TK(u) du, H(\partial_{x_s}p(x))\rangle\\
\overset{s.6}{=}&\frac{\int \|u\|_2^2K(u)du}{D} \langle  I, H(\partial_{x_s}p(x))\rangle\\
=&\frac{\int \|u\|_2^2K(u)du}{D} \Delta(\partial_{x_s}p(x)),
\end{aligned}
\]
where the equation of $s.6$ is obtained from $\|u\|_2^2 = \sum_k u^2_k$, and
\[
\int u_k u_s K(u)du=0 \quad k\neq s,
\] because of symmetric domain of integration and the independence  for each of the dimensions.

Combining the above results, we know the bias
\begin{equation}\label{BiasResult}
|{\rm E}(\partial_{x_s} \hat{p}_h(x)) - \partial_{x_s}p(x)|  = \frac{|\Delta(\partial_{x_s} p(x))|}{2D} h^2 \int \|u\|_2^2 K(u) du+o(h^2),
\end{equation}
where $\Delta(\partial_{x_s} p(x))$ is the Laplace-Beltrami operator of $\partial_{x_s} p(x)$, which is also the summation of the diagonal elements of the Hessian matrix $H(\partial_{x_s}p(x))$. Similarly, repeating the same procedure as \eqref{Expectation}\eqref{Taylor}, we have the second-order bias as
%where $\lambda_s(x)$ is the largest eigenvalue of the Hessian matrix $H(\partial_{x_s}p(x))$. 
\begin{equation}\label{order2}
|{\rm E}( \partial_{x_s} \partial_{x_t}   \hat{p}_h(x)) -  \partial_{x_s} \partial_{x_t} p(x)| = \frac{|\Delta(\partial_{x_s}\partial_{x_t} p(x))|}{2D} h^2 \int \|u\|_2^2 K(u) du+o(h^2).
\end{equation}
The same with \eqref{BiasResult}, $\Delta(\partial_{x_s}\partial_{x_t} p(x))$ is the Laplace-Beltrami operator of $\partial_{x_s}\partial_{x_t} p(x)$ which is also the summation of the eigenvalues of the matrix $M_{s,t}$ whose $i,j$-th element is $\frac{\partial^4}{\partial x_s\partial x_t \partial x_i \partial x_j} p(x)$.%, i.e., $\lambda_{st}(x)= \max_k \lambda_k( M_{st}(x))$.
\end{proof}
\subsection{Derivatives' Variance Bound}\label{variance_proof}
\begin{theorem}
The variance of the first and second order derivatives for $\hat{p}_h(x)$ has a bound as 
\begin{equation*}\label{VarResult}
\begin{aligned}
&{\rm E} |\partial_{x_s} \hat{p}_h(x)-{\rm E}(\partial_{x_s} \hat{p}_h(x))| = \sqrt{\frac{\phi_s(x)}{n h^{D+2}}}+O(\frac{1}{n^{1/2}h^{(D+1)/2}}),\\
&{\rm E} | \partial_{x_s}\partial_{x_t} \hat{p}_h(x) -{\rm E}(\partial_{x_s}\partial_{x_t} \hat{p}_h(x))| = \sqrt{\frac{\phi_{s,t}(x)}{nh^{D+4}}}+O(\frac{1}{n^{1/2}h^{(D+3)/2}}).
\end{aligned}
\end{equation*}
%where $M(u) = \partial_{u_s}K(u)$ and $N(u) = \partial_{u_t}\partial_{u_s}K(u)$, respectively.
\end{theorem}

>>>>

\begin{frontmatter}
\title{supplementary material for ``manifold fitting by ridge estimation: a subspace-constrained approach''}
%\title{A sample article title with some additional note\thanksref{t1}}
\runtitle{Manifold fitting by ridge estimation}
%\thankstext{T1}{A sample additional note to the title.}

@@

\begin{abstract}
In this supplement we present the technical proofs for the main work. Equation and theorem references to the main document do not contain letters.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please use \tableofcontents for articles %%
%% with 50 pages and more                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tableofcontents
\begin{appendix}
\section{PROOFS FOR MAIN THEOREMS AND LEMMAS}
In this section, we prove all key theorems and lemmas in the order in which they appear.
\subsection{Ridge-Derivative Lemma}
\begin{lemma}{ 
For any $R_1$, $R_2$, and any point $x_1\in R_1$, the pairwise distance from $x_1$ to $R_2$ yields the order of %$\min_{x_2\in R_2}\|x_1-x_2\|_2$ % between any two ridges $R_1$ and $R_2$. 
%is of order $O(\|H_1(x_1)-H_2(x_1)\|_F+\|g_1(x_1)-g_2(x_1)\|_2)$, 
\[
\min_{x_2\in R_2}\|x_1-x_2\|_2 = O(\|H_1(x_1)-H_2(x_1)\|_F+\|g_1(x_1)-g_2(x_1)\|_2)
\]
where $H_1(x_1)$ and $g_1 (x_1)$ are the Hessian and gradient of some estimated density function $p_1(x_1)$ evaluated at $x_1$, respectively; $H_2(x_1)$ and $g_2(x_1)$ are the Hessian and gradient of the density function of $p_2(x)$ evaluated at $x_1$, respectively.
%For any two ridges $R, \hat{R}$ defined by $R = \{x| \Pi_{H}(x)g(x) = 0\}$,$\hat{R} = \{x| \Pi_{\hat{H}}(x))\hat{g}(x) = 0\}$, 
 %$d(x, \hat{R}) = O(\|\hat{H}(x)-H(x)\|_F+\|\hat{g}(x)-g(x)\|_2)$
 }\label{margin}
\end{lemma}

The following proof is a revised, simplified version of a similar proof in \cite{genovese2014nonparametric}. For completeness, we include it in our paper.

\begin{proof}\label{Ridge Derivative Lemma}
For two ridges $R_1, R_2$, we have two density functions $p_1(x)$ and $p_2(x)$ such that the points on each ridge satisfy the solution of
 $\Pi_{H_1}(x) g_1(x) = 0$ 
 and $\Pi_{H_2}(x) g_2(x) = 0$, respectively. For any starting point $x_a \in R_1$, we can build a unit speed curve $\gamma(s)$ derived from the gradient and Hessian of $p_2(x)$:
 \begin{gather*}
 \gamma_2(0) = x_a \in R_1,\\
 \gamma_2(t_0) = x_b \in R_2,\\
 \gamma'_2(s) = \frac{\Pi_{H_2}(\gamma(s))g_2(\gamma(s))}{\|\Pi_{H_2}(\gamma(s))g_2(\gamma(s))\|_2}.
 \end{gather*}
Note that the curve $\gamma(t)$ connects $x_a$ with $R_2$ by $x_b$. The univariate function $\xi(s)$ is defined as
  \[
  \xi_2(s) = p_2(\gamma_2(t_0))-p_2(\gamma_2(s)), \quad 0<s<t_0.
  \]
Through a simple computation, we know 
 \[
 \xi'_2(s)=-\langle  g_2(\gamma_2(s)), \gamma'_2(s) \rangle =- \|\Pi_{H_2}(\gamma_2(s))g_2(\gamma_2(s))\|_2, \quad \xi'_2(t_0) = 0.
 \]
 %\[
 %\xi''(s) = \frac{\langle L_{\gamma'(s)}(\gamma(s)) g(\gamma(s)), L(\gamma(s)g(\gamma(s)))\rangle}{\|L(\gamma(s))g(\gamma(s))\|_2}
 %\]
 %with $\gamma(s)=\frac{L(\gamma(s)g(\gamma(s)))}{\|L(\gamma(s)g(\gamma(s)))\|}$
 The distance from $x_a$ to $R_2$ can be bounded by the curve length of $\gamma_2(t)$, which is $t_0$:
 \[
 \begin{aligned}
 &d(x_a, R_2) \\
 =& \|x_a-P_{R_2}(x_a)\|_2\\
 \leq& \|x_a-x_b\|_2 \\
 = &\|\gamma_2(t_0)-\gamma_2(0)\|_2\leq t_0.
 \end{aligned}
 \]
Finally, the problem becomes how best to bound $t_0$. Suppose $\sup_u \xi''_2(u)>\frac{1}{c}$; then, by the mean-value theorem, we have
 \[
 \begin{aligned}
 t_0 = &\frac{\xi'_2(t_0)-\xi'_2(0)}{\xi''_2(u)} \\
 = &\frac{\|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2}{\xi''_2(u)} \\
 \leq & c \|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2.
 \end{aligned}
 \]
 Next, we show that $\|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2$ is of the same order as an approximation error of $H_2(x)$ and $g_2(x)$:
 \begin{equation}
 \begin{aligned}\label{bound_temp1}
 &\|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0))\|_2\\
 =&  \|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0)) - \Pi_{H_1}(\gamma_2(0))g_1(\gamma_2(0))\|_2
 \end{aligned}
 \end{equation}
 Using the triangle inequality with respect to the $\|\|_2$ norm, we determine that \eqref{bound_temp1} is dominated by
 \begin{equation}\label{bound_temp2}
 \begin{aligned}
& \|\Pi_{H_2}(\gamma_2(0))g_2(\gamma_2(0)) -\Pi_{H_1}(\gamma_2(0)){g_2}(\gamma_2(0)) \|_2 +...\\
 & \quad \quad  \quad \quad  \quad \quad + \|\Pi_{H_1}(\gamma_2(0)){g_2}(\gamma_2(0)) - \Pi_{H_1}(\gamma_2(0))g_1(\gamma_2(0))\|_2\\
 \end{aligned}
 \end{equation}
 For any matrix $A$, using the Cauchy-Schwartz inequality on each row of $A$, we will get the result $\|Ax\|_2\leq \|A\|_F\|x\|_2$. Thus, similarly, we determine that \eqref{bound_temp2} is dominated by
 \[
 \begin{aligned}
 & \|\Pi_{H_2}(\gamma_2(0)) - \Pi_{H_1}(\gamma_2(0))\|_F\|g_2(\gamma_2(0))\|_2 +...\\
 &\quad \quad  \quad \quad  \quad \quad  +\|\Pi_{H_1}(\gamma_2(0))\|_F\|g_2(\gamma_2(0))-g_1(\gamma_2(0))\|_2, \\
 \end{aligned}.
 \]
According to the Davis-Kahan theorem, $\|\Pi_{H_2}(\gamma_2(t))-\Pi_{H_1}(\gamma_2(t))\|_F \leq \beta\|H_2(\gamma_2(t))-H_1(\gamma_2(t))\|_F$. The conclusion is therefore proved.
\end{proof}

\subsection{Bound of the Derivatives' Bias}\label{bias_proof}
\begin{theorem}
The bias of the first order and second order of the $\hat{p}_h(x)$ is 
\begin{equation*}\label{BiasResult}
|{\rm E}(\partial_{x_s}  \hat{p}_h(x)) - \partial_{x_s}p(x)|  = \frac{h^2|\Delta (\partial_{x_s}p(x))|}{2D}  \int \|u\|_2^2 K(u) du+o(h^2),
\end{equation*}
\begin{equation*}\label{order2}
|{\rm E}( \partial_{x_s} \partial_{x_t}  \hat{p}_h(x)) -  \partial_{x_s} \partial_{x_t} p(x)| = \frac{h^2 |\Delta (\partial_{x_s}\partial_{x_t} p(x))|}{2D} \int \|u\|_2^2 K(u) du+o(h^2),
\end{equation*}
where $\Delta$ is the Laplace-Beltrami operator.
\end{theorem}

\begin{proof}
Suppose the kernel function vanishes at infinity for each dimension, i.e., it satisfies $\lim_{u_s\rightarrow \infty}K(u)=0$ for each dimension. Then, using the integration-by-parts formula, we obtain the expectation of first-order derivatives:
\begin{equation}\label{Expectation}
\begin{aligned}
&{\rm E} (\partial_{x_s} \hat{p}_h(x)) \\
\overset{s.1}{=}& \frac{1}{h^D} \int_{y\in \mathbb{R}^D}{\partial_{x_s}}  K(\frac{x-y}{h}) p(y) dy  \\
\overset{s.2}{=}&\frac{1}{h^{D+1}}\int \partial_{z_s} K(z)|_{z=\frac{x-y}{h}} p(y) dy\\
\overset{s.3}{=}&h^{-1} \int_ {u\in \mathbb{R}^D} \partial_{u_s} K(u) p(x-hu)  d u\\
\overset{s.4}{=}&h^{-1}\int_ {u\in \mathbb{R}^D} K(u) \partial_{u_s} p(x-hu) d u \\
\overset{s.5}{=}& \int_ {u\in \mathbb{R}^D} K(u) \partial_{z_s}p(z)|_{z=x-hu} d u.
\end{aligned}
\end{equation}

The equation $s.1$ is the definition of the expectation. The equation $s.2$ is obtained from the derivative of the function composition. The equation $s.3$ is obtained from the formula of the integration by changing variables. The equation $s.4$ is obtained from the formula of the integration by partition. The equation $s.5$ is similar to equation $s.2$.

@@

For the term $u^T H(\partial_{x_s}p(x))u$, we know it is related to the Laplace-Beltrami operator of $\Delta(\partial_{x_s}p(x))$ by
\[
\begin{aligned}
\int u^T H(\partial_{x_s}p(x))u K(u) du =& \langle \int uu^TK(u) du, H(\partial_{x_s}p(x))\rangle\\
\overset{s.6}{=}&\frac{\int \|u\|_2^2K(u)du}{D} \langle  I, H(\partial_{x_s}p(x))\rangle\\
=&\frac{\int \|u\|_2^2K(u)du}{D} \Delta(\partial_{x_s}p(x)),
\end{aligned}
\]
where the equation of $s.6$ is obtained from $\|u\|_2^2 = \sum_k u^2_k$, and
\[
\int u_k u_s K(u)du=0 \quad k\neq s,
\] because of symmetric domain of integration and the independence  of each of the dimensions.
